Reproducible Research: Peer Assessment 1
========================================================

This assignment will be described in multiple parts.It will write a report that answers the questions detailed below:

+ What is mean total number of steps taken per day?
+ What is the average daily activity pattern?
+ Imputing missing values
+ Are there differences in activity patterns between weekdays and weekends?

Assignment
----------

### Loading and preprocessing the data

```{r}
echo = TRUE  # Always make code visible
options(scipen = 1)  # Turn off scientific notations for numbers
```

#### Reformat and remove NA values

```{r}
library(lattice)
library(ggplot2)
activity.data <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
activity.data$date <- as.Date(activity.data$date, "%Y-%m-%d")
head(activity.data)
names(activity.data)
activity.data.noNA <- na.omit(activity.data)
head(activity.data.noNA)

```

### What is mean total number of steps taken per day?

#### 1. Total number of steps taken per day

```{r}
Total.Steps.per.Day <- aggregate(steps ~ date, data = activity.data.noNA, sum)
Total.Steps.per.Day

```

#### 2. Histogram of the total number of steps taken each day

```{r}
ggplot(activity.data.noNA, aes(date, steps)) + geom_bar(stat = "identity", colour = "#00C0AF", fill = "#00C0AF", width = .6) + labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")

```

#### 3. Mean and median of the total number of steps taken per day

Mean:
```{r}
Total.Steps <- aggregate(activity.data.noNA$steps, list(Date = activity.data.noNA$date), FUN = "sum")$x
mean(Total.Steps)

```

Median:
```{r}
median(Total.Steps)
```

### What is the average daily activity pattern?

#### 1. Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
Time.Series <- tapply(activity.data.noNA$steps, activity.data.noNA$interval, mean)
plot(row.names(Time.Series), Time.Series, type = "l", xlab = "5-minute interval", ylab = "Average across all Days", main = "Average number of steps taken",     col = "#E58700")

```

#### 2. 5-minute interval, on average across all the days in the dataset, having the maximum number of steps

```{r}
Max.Interval <- which.max(Time.Series)
names(Max.Interval)

```

### Imputing missing values

#### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
NA.Activities <- sum(is.na(activity.data))
NA.Activities
```

#### 2. Fill in all of the missing values in the dataset with mean for that day
```{r}
Steps.Mean <- aggregate(steps ~ interval, data = activity.data, FUN = mean)
fill.NA <- numeric()
for (i in 1:nrow(activity.data)) {
    activity.data2 <- activity.data[i, ]
    if (is.na(activity.data2$steps)) {
        steps <- subset(Steps.Mean, interval == activity.data2$interval)$steps
    } else {
        steps <- activity.data2$steps
    }
    fill.NA <- c(fill.NA, steps)
}
```

#### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
new.activity.data <- activity.data
new.activity.data$steps <- fill.NA
```


#### 4. Histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r}
ggplot(new.activity.data, aes(date, steps)) + geom_bar(stat = "identity",
                                             colour = "#B983FF",
                                             fill = "#B983FF",
                                             width = 0.6) + labs(title = "Histogram of Total Number of Steps Taken Each Day (no missing data)", x = "Date", y = "Total number of steps")

```

#### Do these values differ from the estimates from the first part of the assignment?

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?


```{r}
New.Total.Steps <- aggregate(new.activity.data$steps, 
                           list(Date = new.activity.data$date), 
                           FUN = "sum")$x
New.Mean <- mean(New.Total.Steps)
New.Mean

New.Median <- median(New.Total.Steps)
New.Median

Old.Mean <- mean(Total.Steps)
Old.Median <- median(Total.Steps)
New.Mean - Old.Mean

New.Median - Old.Median
```

After replacing the mean is the same but the median is a little bit different


### Are there differences in activity patterns between weekdays and weekends?

#### 1. Create a new factor variable to in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day

```{r}
new.activity.data$weekDays <- factor(format(new.activity.data$date, "%A"))
levels(new.activity.data$weekDays)

levels(new.activity.data$weekDays) <- list(weekday = c("Monday", "Tuesday",                                      "Wednesday", "Thursday", "Friday"), weekend = c("Saturday", "Sunday"))

levels(new.activity.data$weekDays)

```

#### 2. A panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
Steps.By.Days <- aggregate(steps ~ interval + weekDays, data = new.activity.data, mean)

names(Steps.By.Days) <- c("interval", "daylevel", "steps")
xyplot(steps ~ interval | daylevel, Steps.By.Days, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```
